Symbolic Abstract Data Type Inference

### 2015-11-02 draft.tex
Final version candidate.

\bibliographystyle{abbrvnat}
\bibliography{violin}

% % The bibliography should be embedded for final submission.
%
% \begin{thebibliography}{}
% \softraggedright
%
% \bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
% P. Q. Smith, and X. Y. Jones. ...reference text...
%
% \end{thebibliography}

>>>>

\bibliographystyle{abbrvnat}

% \bibliography{dblp,books}

\begin{thebibliography}{26}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Abdulla et~al.(1996)Abdulla, Cerans, Jonsson, and
  Tsay]{conf/lics/AbdullaCJT96}
P.~A. Abdulla, K.~Cerans, B.~Jonsson, and Y.~Tsay.
\newblock General decidability theorems for infinite-state systems.
\newblock In \emph{Proceedings, 11th Annual {IEEE} Symposium on Logic in
  Computer Science, New Brunswick, New Jersey, USA, July 27-30, 1996}, pages
  313--321. {IEEE} Computer Society, 1996.
\newblock \url{http://dx.doi.org/10.1109/LICS.1996.561359}.

\bibitem[Abdulla et~al.(2013)Abdulla, Haziza, Hol{\'{\i}}k, Jonsson, and
  Rezine]{conf/tacas/AbdullaHHJR13}
P.~A. Abdulla, F.~Haziza, L.~Hol{\'{\i}}k, B.~Jonsson, and A.~Rezine.
\newblock An integrated specification and verification technique for highly
  concurrent data structures.
\newblock In \emph{Tools and Algorithms for the Construction and Analysis of
  Systems - 19th International Conference, {TACAS} 2013, Held as Part of the
  European Joint Conferences on Theory and Practice of Software, {ETAPS} 2013,
  Rome, Italy, March 16-24, 2013. Proceedings}, volume 7795 of \emph{Lecture
  Notes in Computer Science}, pages 324--338. Springer, 2013.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-36742-7_23}.

\bibitem[Amit et~al.(2007)Amit, Rinetzky, Reps, Sagiv, and
  Yahav]{conf/cav/AmitRRSY07}
D.~Amit, N.~Rinetzky, T.~W. Reps, M.~Sagiv, and E.~Yahav.
\newblock Comparison under abstraction for verifying linearizability.
\newblock In \emph{Computer Aided Verification, 19th International Conference,
  {CAV} 2007, Berlin, Germany, July 3-7, 2007, Proceedings}, volume 4590 of
  \emph{Lecture Notes in Computer Science}, pages 477--490. Springer, 2007.
\newblock \url{http://dx.doi.org/10.1007/978-3-540-73368-3_49}.

\bibitem[Angluin(1987)]{journals/iandc/Angluin87}
D.~Angluin.
\newblock Learning regular sets from queries and counterexamples.
\newblock \emph{Inf. Comput.}, 75\penalty0 (2):\penalty0 87--106, 1987.
\newblock \url{http://dx.doi.org/10.1016/0890-5401(87)90052-6}.

\bibitem[Bouajjani et~al.(2015{\natexlab{a}})Bouajjani, Emmi, Enea, and
  Hamza]{conf/icalp/BouajjaniEEH15}
A.~Bouajjani, M.~Emmi, C.~Enea, and J.~Hamza.
\newblock On reducing linearizability to state reachability.
\newblock In \emph{Automata, Languages, and Programming - 42nd International
  Colloquium, {ICALP} 2015, Kyoto, Japan, July 6-10, 2015, Proceedings, Part
  {II}}, volume 9135 of \emph{Lecture Notes in Computer Science}, pages
  95--107. Springer, 2015{\natexlab{a}}.
\newblock \url{http://dx.doi.org/10.1007/978-3-662-47666-6_8}.

\bibitem[Bouajjani et~al.(2015{\natexlab{b}})Bouajjani, Emmi, Enea, and
  Hamza]{conf/popl/BouajjaniEEH15}
A.~Bouajjani, M.~Emmi, C.~Enea, and J.~Hamza.
\newblock Tractable refinement checking for concurrent objects.
\newblock In \emph{Proceedings of the 42nd Annual {ACM} {SIGPLAN-SIGACT}
  Symposium on Principles of Programming Languages, {POPL} 2015, Mumbai, India,
  January 15-17, 2015}, pages 651--662. {ACM}, 2015{\natexlab{b}}.
\newblock \url{http://doi.acm.org/10.1145/2676726.2677002}.

\bibitem[Burckhardt et~al.(2010)Burckhardt, Dern, Musuvathi, and
  Tan]{conf/pldi/BurckhardtDMT10}
S.~Burckhardt, C.~Dern, M.~Musuvathi, and R.~Tan.
\newblock Line-up: a complete and automatic linearizability checker.
\newblock In \emph{Proceedings of the 2010 {ACM} {SIGPLAN} Conference on
  Programming Language Design and Implementation, {PLDI} 2010, Toronto,
  Ontario, Canada, June 5-10, 2010}, pages 330--340. {ACM}, 2010.
\newblock \url{http://doi.acm.org/10.1145/1806596.1806634}.

\bibitem[Burnim et~al.(2011)Burnim, Necula, and Sen]{conf/asplos/BurnimNS11}
J.~Burnim, G.~C. Necula, and K.~Sen.
\newblock Specifying and checking semantic atomicity for multithreaded
  programs.
\newblock In \emph{Proceedings of the 16th International Conference on
  Architectural Support for Programming Languages and Operating Systems,
  {ASPLOS} 2011, Newport Beach, CA, USA, March 5-11, 2011}, pages 79--90.
  {ACM}, 2011.
\newblock \url{http://doi.acm.org/10.1145/1950365.1950377}.

\bibitem[Dodds et~al.(2015)Dodds, Haas, and Kirsch]{conf/popl/DoddsHK15}
M.~Dodds, A.~Haas, and C.~M. Kirsch.
\newblock A scalable, correct time-stamped stack.
\newblock In \emph{Proceedings of the 42nd Annual {ACM} {SIGPLAN-SIGACT}
  Symposium on Principles of Programming Languages, {POPL} 2015, Mumbai, India,
  January 15-17, 2015}, pages 233--246. {ACM}, 2015.
\newblock \url{http://doi.acm.org/10.1145/2676726.2676963}.

\bibitem[Dragoi et~al.(2013)Dragoi, Gupta, and Henzinger]{conf/cav/DragoiGH13}
C.~Dragoi, A.~Gupta, and T.~A. Henzinger.
\newblock Automatic linearizability proofs of concurrent objects with
  cooperating updates.
\newblock In \emph{Computer Aided Verification - 25th International Conference,
  {CAV} 2013, Saint Petersburg, Russia, July 13-19, 2013. Proceedings}, volume
  8044 of \emph{Lecture Notes in Computer Science}, pages 174--190. Springer,
  2013.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-39799-8_11}.

\bibitem[Emmi et~al.(2015)Emmi, Enea, and Hamza]{conf/pldi/EmmiEH15}
M.~Emmi, C.~Enea, and J.~Hamza.
\newblock Monitoring refinement via symbolic reasoning.
\newblock In \emph{Proceedings of the 36th {ACM} {SIGPLAN} Conference on
  Programming Language Design and Implementation, Portland, OR, USA, June
  15-17, 2015}, pages 260--269. {ACM}, 2015.
\newblock \url{http://doi.acm.org/10.1145/2737924.2737983}.

\bibitem[Finkel and Schnoebelen(2001)]{journals/tcs/FinkelS01}
A.~Finkel and P.~Schnoebelen.
\newblock Well-structured transition systems everywhere!
\newblock \emph{Theor. Comput. Sci.}, 256\penalty0 (1-2):\penalty0 63--92,
  2001.
\newblock \url{http://dx.doi.org/10.1016/S0304-3975(00)00102-X}.

\bibitem[Fishburn(1985)]{books/Fishburn85}
P.~C. Fishburn.
\newblock \emph{Interval Orders and Interval Graphs: A Study of Partially
  Ordered Sets}.
\newblock Wiley-Interscience Series in Discrete Mathematics. John Wiley \& Sons
  Inc, 1985.

\bibitem[Giannakopoulou et~al.(2005)Giannakopoulou, Pasareanu, and
  Barringer]{journals/ase/GiannakopoulouPB05}
D.~Giannakopoulou, C.~S. Pasareanu, and H.~Barringer.
\newblock Component verification with automatically generated assumptions.
\newblock \emph{Autom. Softw. Eng.}, 12\penalty0 (3):\penalty0 297--320, 2005.
\newblock \url{http://dx.doi.org/10.1007/s10515-005-2641-y}.

\bibitem[Hemed and Rinetzky(2014)]{conf/podc/HemedR14}
N.~Hemed and N.~Rinetzky.
\newblock Brief announcement: concurrency-aware linearizability.
\newblock In \emph{{ACM} Symposium on Principles of Distributed Computing,
  {PODC} '14, Paris, France, July 15-18, 2014}, pages 209--211. {ACM}, 2014.
\newblock \url{http://doi.acm.org/10.1145/2611462.2611513}.

\bibitem[Hendler et~al.(2006)Hendler, Lev, Moir, and
  Shavit]{journals/dc/HendlerLMS06}
D.~Hendler, Y.~Lev, M.~Moir, and N.~Shavit.
\newblock A dynamic-sized nonblocking work stealing deque.
\newblock \emph{Distributed Computing}, 18\penalty0 (3):\penalty0 189--207,
  2006.
\newblock \url{http://dx.doi.org/10.1007/s00446-005-0144-5}.

\bibitem[Henzinger et~al.(2013)Henzinger, Sezgin, and
  Vafeiadis]{conf/concur/HenzingerSV13}
T.~A. Henzinger, A.~Sezgin, and V.~Vafeiadis.
\newblock Aspect-oriented linearizability proofs.
\newblock In \emph{{CONCUR} 2013 - Concurrency Theory - 24th International
  Conference, {CONCUR} 2013, Buenos Aires, Argentina, August 27-30, 2013.
  Proceedings}, volume 8052 of \emph{Lecture Notes in Computer Science}, pages
  242--256. Springer, 2013.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-40184-8_18}.

\bibitem[Herlihy and Wing(1990)]{journals/toplas/HerlihyW90}
M.~Herlihy and J.~M. Wing.
\newblock Linearizability: {A} correctness condition for concurrent objects.
\newblock \emph{{ACM} Trans. Program. Lang. Syst.}, 12\penalty0 (3):\penalty0
  463--492, 1990.
\newblock \url{http://doi.acm.org/10.1145/78969.78972}.

\bibitem[Liang and Feng(2013)]{conf/pldi/LiangF13}
H.~Liang and X.~Feng.
\newblock Modular verification of linearizability with non-fixed linearization
  points.
\newblock In \emph{{ACM} {SIGPLAN} Conference on Programming Language Design
  and Implementation, {PLDI} '13, Seattle, WA, USA, June 16-19, 2013}, pages
  459--470. {ACM}, 2013.
\newblock \url{http://doi.acm.org/10.1145/2462156.2462189}.

\bibitem[Liu et~al.(2009)Liu, Chen, Liu, and Sun]{conf/fm/LiuCLS09}
Y.~Liu, W.~Chen, Y.~A. Liu, and J.~Sun.
\newblock Model checking linearizability via refinement.
\newblock In \emph{{FM} 2009: Formal Methods, Second World Congress, Eindhoven,
  The Netherlands, November 2-6, 2009. Proceedings}, volume 5850 of
  \emph{Lecture Notes in Computer Science}, pages 321--337. Springer, 2009.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-05089-3_21}.

\bibitem[O'Hearn et~al.(2010)O'Hearn, Rinetzky, Vechev, Yahav, and
  Yorsh]{conf/podc/OHearnRVYY10}
P.~W. O'Hearn, N.~Rinetzky, M.~T. Vechev, E.~Yahav, and G.~Yorsh.
\newblock Verifying linearizability with hindsight.
\newblock In \emph{Proceedings of the 29th Annual {ACM} Symposium on Principles
  of Distributed Computing, {PODC} 2010, Zurich, Switzerland, July 25-28,
  2010}, pages 85--94. {ACM}, 2010.
\newblock \url{http://doi.acm.org/10.1145/1835698.1835722}.

\bibitem[Shacham et~al.(2011)Shacham, Bronson, Aiken, Sagiv, Vechev, and
  Yahav]{conf/oopsla/ShachamBASVY11}
O.~Shacham, N.~G. Bronson, A.~Aiken, M.~Sagiv, M.~T. Vechev, and E.~Yahav.
\newblock Testing atomicity of composed concurrent operations.
\newblock In \emph{Proceedings of the 26th Annual {ACM} {SIGPLAN} Conference on
  Object-Oriented Programming, Systems, Languages, and Applications, {OOPSLA}
  2011, part of {SPLASH} 2011, Portland, OR, USA, October 22 - 27, 2011}, pages
  51--64. {ACM}, 2011.
\newblock \url{http://doi.acm.org/10.1145/2048066.2048073}.

\bibitem[Vafeiadis(2010)]{conf/cav/Vafeiadis10}
V.~Vafeiadis.
\newblock Automatically proving linearizability.
\newblock In \emph{Computer Aided Verification, 22nd International Conference,
  {CAV} 2010, Edinburgh, UK, July 15-19, 2010. Proceedings}, volume 6174 of
  \emph{Lecture Notes in Computer Science}, pages 450--464. Springer, 2010.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-14295-6_40}.

\bibitem[Wing and Gong(1993)]{journals/jpdc/WingG93}
J.~M. Wing and C.~Gong.
\newblock Testing and verifying concurrent objects.
\newblock \emph{J. Parallel Distrib. Comput.}, 17\penalty0 (1-2):\penalty0
  164--182, 1993.
\newblock \url{http://dx.doi.org/10.1006/jpdc.1993.1015}.

\bibitem[Zhang et~al.(2013)Zhang, Chattopadhyay, and Wang]{conf/kbse/ZhangCW13}
L.~Zhang, A.~Chattopadhyay, and C.~Wang.
\newblock Round-up: Runtime checking quasi linearizability of concurrent data
  structures.
\newblock In \emph{2013 28th {IEEE/ACM} International Conference on Automated
  Software Engineering, {ASE} 2013, Silicon Valley, CA, USA, November 11-15,
  2013}, pages 4--14. {IEEE}, 2013.
\newblock \url{http://dx.doi.org/10.1109/ASE.2013.6693061}.

\bibitem[Zhang(2011)]{conf/icse/Zhang11a}
S.~J. Zhang.
\newblock Scalable automatic linearizability checking.
\newblock In \emph{Proceedings of the 33rd International Conference on Software
  Engineering, {ICSE} 2011, Waikiki, Honolulu , HI, USA, May 21-28, 2011},
  pages 1185--1187. {ACM}, 2011.
\newblock \url{http://doi.acm.org/10.1145/1985793.1986037}.

\end{thebibliography}

### 2015-11-02 related.tex
Final version candidate.

The idea of decomposing ADT violations into a finite set of patterns has been
proposed for atomic collections~\cite{conf/tacas/AbdullaHHJR13,
conf/concur/HenzingerSV13, conf/popl/DoddsHK15, conf/popl/BouajjaniEEH15}, and
generalized to ADTs which can be expressed with a particular form of recursive
definition~\cite{journals/arxiv/BouajjaniEEH15}. These works suggest patterns
which directly recognize violations in \emph{concurrent} executions,
effectively reducing observational refinement to the verification of classical
temporal-logic properties. Some of these patterns are expressible only as
\emph{infinite}, though regular, languages. On the contrary, in this work we
suggest patterns expressed as \emph{finite} languages, which recognize
violations in ADTs — i.e.,~the sequential executions of atomic
objects, or the “concurrency-aware” executions of “concurrency-aware”
specifications~\cite{conf/podc/HemedR14}. Though by themselves these patterns
are incomplete in recognizing violations in concurrent executions, they are
complete when considered as negative conditions on the linearizations of
concurrent executions, effectively reducing verification of a single execution
to satisfiability checking~\cite{conf/pldi/EmmiEH15}.

>>>>

The idea of decomposing ADT violations into a finite set of patterns has been
proposed for atomic collections~\cite{conf/tacas/AbdullaHHJR13,
conf/concur/HenzingerSV13, conf/popl/DoddsHK15, conf/popl/BouajjaniEEH15}, and
generalized to ADTs which can be expressed with a particular form of recursive
definition~\cite{conf/icalp/BouajjaniEEH15}. These works suggest patterns
which directly recognize violations in \emph{concurrent} executions,
effectively reducing observational refinement to the verification of classical
temporal-logic properties. Some of these patterns are expressible only as
\emph{infinite}, though regular, languages. On the contrary, in this work we
suggest patterns expressed as \emph{finite} languages, which recognize
violations in ADTs — i.e.,~the sequential executions of atomic
objects, or the “concurrency-aware” executions of “concurrency-aware”
specifications~\cite{conf/podc/HemedR14}. Though by themselves these patterns
are incomplete in recognizing violations in concurrent executions, they are
complete when considered as negative conditions on the linearizations of
concurrent executions, effectively reducing verification of a single execution
to satisfiability checking~\cite{conf/pldi/EmmiEH15}.
